---
title: "Resuls of MCS for 802.11ax and 802.11n"
author: "Bruno Sousa"
date: "6/12/2020"
output:
  html_document: default
  word_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Some tests have failed to run. In particular the test 06 which gives an error, "No free association ID available!".

The association ID varies between 1 and 2007. In the scenario we have 400 nodes, So this should not happen, at least based on the number of nodes in the scenario. 
Additionally, the remaining scenarios should also have this issue, in particular the test_06

This happens in the beginning of the simulation, that is when stations are connecting to the APs!

```{r echo=TRUE}
ExpID_to_ignore <- c(6)
```



```{r init, echo=FALSE}
library(plotly)
library(data.table)
library(ggplot2)

# Script to check the results
#

# MErge with excel settings
version<- 1.1
load(file= paste0("resdata_", version, ".data"))
dataExperiments <- data.table(read.csv2(paste0(dir_base, '/Metricsv2.csv'), header=TRUE, sep=';') )

data_signalDBM$ExpID <- as.numeric(get_id_from_str(data_signalDBM$experiment))
data_noiseDBM$ExpID <- as.numeric(get_id_from_str(data_noiseDBM$experiment))
data_remainEnergy$ExpID <- as.numeric(get_id_from_str(data_remainEnergy$experiment))
data_consumedEnergy$ExpID <- as.numeric(get_id_from_str(data_consumedEnergy$experiment))
data_flow$ExpID <- as.numeric(get_id_from_str(data_flow$experiment))


dt_signalDBM <- merge.data.table(dataExperiments, data_signalDBM, by.x=c('ExpID'), by.y=c('ExpID'))
dt_noiseDBM <- merge.data.table(dataExperiments, data_noiseDBM, by.x=c('ExpID'), by.y=c('ExpID'))
dt_remainingEnergy <- merge.data.table(dataExperiments, data_remainEnergy, by.x=c('ExpID'), by.y=c('ExpID'))
dt_consumedEnergy  <- merge.data.table(dataExperiments, data_consumedEnergy, by.x=c('ExpID'), by.y=c('ExpID'))
dt_flow <- merge.data.table(dataExperiments, data_flow, by.x=c('ExpID'), by.y=c('ExpID'))
```

# Signal DBM

**Higher values of dBm** mean that more power is required to transmit or receive.
(class SignalNoiseDbm) in Wifi-Phy.h file.

See https://www.networkcomputing.com/wireless-infrastructure/wifi-networking-radio-wave-basics

```{r signalDBM, echo=FALSE}
g <- ggplot(data=dt_signalDBM, aes(x=factor(ExpID), y=Mean, fill=factor(X..technology), colour=factor(X..numSta)) )
g <- g + geom_boxplot()
g <- g + xlab(' Signal DBM per technology (0=ax, 1=n,) and number of Stations ')
fig  <- ggplotly(g)
fig
```

**Results**

* Tech: The 802.11ax has more variation of signal.
* N. of Clients: The increased number of clients leads to lower the variation in both technologies.



```{r signalDBM_frequency, echo=FALSE}
g <- ggplot(data=dt_signalDBM[!ExpID %in% ExpID_to_ignore,], aes(x=factor(X..frequency), y=Mean, fill=factor(X..frequency) ))
g <- g + geom_boxplot()
g <- g + xlab(' Signal DBM per Band frequency (igoring bad tests) ')
fig1  <- ggplotly(g)
fig1
```

**Results**

* Band: The 6GHz introduces higher variations in the signal and higher values for signal (more power)
* 



```{r signalDBM_technology_channelWidth, echo=FALSE}
dt_signalDBM$tech_channel <- paste0(dt_signalDBM$X..technology, "_", dt_signalDBM$X..channelWidth)
g <- ggplot(data=dt_signalDBM[ !ExpID %in% ExpID_to_ignore, ], aes(x=factor(tech_channel), y=Mean, fill=factor(tech_channel) ))
g <- g + geom_boxplot()
g <- g + xlab(' Signal DBM per technology (0=ax, 1=n,) and channelBandwidth (igoring bad tests) ')
fig1  <- ggplotly(g)
fig1
```


**Results**

* Channel Width: 
* 





# Noise

**Higher values of noise dBm** mean that more the signal may become imperceptible, thus not good.
(class SignalNoiseDbm) in Wifi-Phy.h file.

See https://www.networkcomputing.com/wireless-infrastructure/wifi-networking-radio-wave-basics

See https://www.cisco.com/en/US/docs/solutions/Enterprise/Mobility/emob30dg/RFDesign.html#wp999557 

```{r noiseDBM, echo=FALSE}
g <- ggplot(data=dt_noiseDBM, aes(x=factor(ExpID), y=Mean, fill=factor(X..technology), colour=factor(X..numSta)) )
g <- g + geom_boxplot()
g <- g + xlab(' Noise DBM per technology (0=ax, 1=n,) and number of Stations ')
fig  <- ggplotly(g)
fig
```

**Results**

* Tech: 802.11ax has lower noise in the band 2Ghz, in the 5Ghz and 6GHz it does not have this behaviour. The difference between the tests is marginal.
* N. of Clients: The noise tends to increase with the number of clients.



```{r noiseDBM_frequency, echo=FALSE}
g <- ggplot(data=dt_noiseDBM[!ExpID %in% ExpID_to_ignore,], aes(x=factor(X..frequency), y=Mean, fill=factor(X..frequency) ))
g <- g + geom_boxplot()
g <- g + xlab(' Noise DBM per Band frequency (ignoring tests)')
fig1  <- ggplotly(g)
fig1
```

**Results**

* Band: the 5GHz and 6GHz lead to lower variation of noise and lower levels of noise.
* 



```{r noiseDBM_technology_channelWidth, echo=FALSE}
dt_noiseDBM$tech_channel <- paste0(dt_noiseDBM$X..technology, "_", dt_noiseDBM$X..channelWidth)
g <- ggplot(data=dt_noiseDBM[!ExpID %in% ExpID_to_ignore, ], aes(x=factor(tech_channel), y=Mean, fill=factor(tech_channel) ))
g <- g + geom_boxplot()
g <- g + xlab(' Noise DBM per technology (0=ax, 1=n,) and channelBandwidth (ignoring test)')
fig1  <- ggplotly(g)
fig1
```


**Results**

* Channel Width: 
* 




## Remaining Energy 

*Initial values of energy* are



```{r energyRemain, echo=FALSE}
#
# DT Stations
#
# 1 - count NAs per Experiment
dt_eneRemain <- data.table(dt_remainingEnergy)
#dt_remainingEnergy_NAs <- dt_STA[is.na(Median) , list(N=N, Mean=Mean, Median=Median, SD=SD), by=c("ExpID","nodeId") ]
dt_STA_NAs <- dt_eneRemain[ , list(Na=sum(is.na(nodeId)) ), by=c("ExpID") ]
dt_STA_NAs <- merge.data.table(dataExperiments, dt_STA_NAs, by.x=c('ExpID'), by.y=c('ExpID'))

g <- ggplot(data=dt_STA_NAs, aes(x=factor(ExpID), y=Na, fill=factor(X..frequency) ))
g <- g + geom_bar(position="dodge", stat="identity")
g <- g + xlab(" Number of NAs in the tests ")
fig2a  <- ggplotly(g)
fig2a
```
### Remaining Energy per Technology and test

```{r energyPerTech, echo=FALSE}
g <- ggplot(data=dt_remainingEnergy, aes(x=factor(ExpID), y=Mean, fill=factor(X..technology), colour=factor(X..numSta) ))
#g <- g + geom_bar(position="dodge", stat="identity")
g <- g + geom_boxplot()
g <- g + ylab('Remaining Energy (higher values better)')
g <- g + xlab('Technology and number of clients')
fig2a  <- ggplotly(g)
fig2a
```
**Results:**

* Tech: 802.11ax has lower values for the remaining energy, which leads to lower performance. The band 5GHz seems to be more efficient in terms of Remaining energy 
* N. of Clients: The different is quite neglactable, but the higher number of clients tends to lead to lower remaining energy levels.


```{r energyRemainPerTechBand, echo=FALSE}
g <- ggplot(data=dt_remainingEnergy[!ExpID %in% ExpID_to_ignore,], aes(x=factor(X..frequency), y=Mean, fill=factor(X..frequency) ))
g <- g + geom_boxplot()
g <- g + ylab('Remaining Energy (higher values better)')
g <- g + xlab('Band ')
fig1  <- ggplotly(g)
fig1
```

**Results:**

* Band: the 5and 6GHz tend to lead to best levels of remaining energy





```{r remainingEnergy_channelWidth, echo=FALSE}
dt_remainingEnergy$tech_channel <- paste0(dt_remainingEnergy$X..technology, "_", dt_remainingEnergy$X..channelWidth)
g <- ggplot(data=dt_remainingEnergy[!ExpID %in% ExpID_to_ignore, ], aes(x=factor(tech_channel), y=Mean, fill=factor(tech_channel) ))
g <- g + geom_boxplot()
g <- g + ylab('Remaining Energy (higher values better)')
g <- g + xlab('Per technology (0=ax, 1=n,) and channelBandwidth (ignoring test)')
fig1  <- ggplotly(g)
fig1
```


**Results**

* Channel Width: 802.11ax tends to be more "stable"
* 







## Consumed Energy 

*Initial levels of energy* are




### Consumed Energy per Technology and test

```{r energyConsumedPerTech, echo=FALSE}
g <- ggplot(data=dt_consumedEnergy, aes(x=factor(ExpID), y=Mean, fill=factor(X..technology), colour=factor(X..numSta) ))
#g <- g + geom_bar(position="dodge", stat="identity")
g <- g + geom_boxplot()
g <- g + ylab('Consumed Energy (lower values better)')
g <- g + xlab('Technology and number of clients')
fig2a  <- ggplotly(g)
fig2a
```
**Results:**

* Tech: 802.11ax has higher values for the consumed energy, which leads to lower performance. The band 5GHz seems to be more efficient in terms of Remaining energy 
* N. of Clients: The different is quite neglactable, but the higher number of clients tends to lead to lower remaining energy levels.


```{r energyConsumedPerTechBand, echo=FALSE}
g <- ggplot(data=dt_consumedEnergy[!ExpID %in% ExpID_to_ignore,], aes(x=factor(X..frequency), y=Mean, fill=factor(X..frequency) ))
g <- g + geom_boxplot()
g <- g + ylab('Remaining Energy (lower values better)')
g <- g + xlab('Band ')
fig1  <- ggplotly(g)
fig1
```

**Results:**

* Band: the 5and 6GHz tend to lead to best levels of consumed energy





```{r consumedEnergy_channelWidth, echo=FALSE}
dt_consumedEnergy$tech_channel <- paste0(dt_consumedEnergy$X..technology, "_", dt_consumedEnergy$X..channelWidth)
g <- ggplot(data=dt_consumedEnergy[!ExpID %in% ExpID_to_ignore, ], aes(x=factor(tech_channel), y=Mean, fill=factor(tech_channel) ))
g <- g + geom_boxplot()
g <- g + ylab('Consumed Energy (lower values better)')
g <- g + xlab('Per technology (0=ax, 1=n,) and channelBandwidth (ignoring test)')
fig1  <- ggplotly(g)
fig1
```


**Results**

* Channel Width: 802.11ax tends to be more "stable" for the 20 MHz channel.
* 


# Flow information

## Delay 


```{r dt_flow, echo=FALSE}

g <- ggplot(data=dt_flow, aes(x=factor(FlowID), y=MeanDelay, fill=factor(X..technology), colour=factor(X..numSta) ))
#g <- g + geom_bar(position="dodge", stat="identity")
g <- g + geom_boxplot()
g <- g + ylab('MEan Delay per Experiment')
g <- g + xlab('Technology and number of clients within flowID and experiments')
g <- g + facet_wrap(~ExpID, scales="free")
fig2a  <- ggplotly(g)
fig2a
```

**Results:**

* Tech:  
* N. of Clients: 


```{r FlowPerTechBand, echo=FALSE}
g <- ggplot(data=dt_flow[!ExpID %in% ExpID_to_ignore,], aes(x=factor(X..frequency), y=MeanDelay, fill=factor(X..frequency) ))
g <- g + geom_boxplot()
g <- g + ylab('Mean Delay')
g <- g + xlab('Band ')
fig1  <- ggplotly(g)
fig1
```


```{r flow_channelWidth, echo=FALSE}
dt_flow$tech_channel <- paste0(dt_flow$X..technology, "_", dt_flow$X..channelWidth)
g <- ggplot(data=dt_flow[!ExpID %in% ExpID_to_ignore, ], aes(x=factor(tech_channel), y=MeanDelay, fill=factor(tech_channel) ))
g <- g + geom_boxplot()
g <- g + ylab('Mean Delay')
g <- g + xlab('Per technology (0=ax, 1=n,) and channelBandwidth (ignoring test)')
fig1  <- ggplotly(g)
fig1
```


## TXBitRate

```{r dt_flow_TX, echo=FALSE}

g <- ggplot(data=dt_flow, aes(x=factor(TXBitrate), y=MeanDelay, fill=factor(X..technology), colour=factor(X..numSta) ))
g <- g + geom_boxplot()
g <- g + ylab('TXBitrate per Experiment')
g <- g + xlab('Technology and number of clients within flowID and experiments')
g <- g + facet_wrap(~ExpID, scales="free")
fig2a  <- ggplotly(g)
fig2a
```

**Results:**

* Tech:  
* N. of Clients: 


```{r Flow_TX, echo=FALSE}
g <- ggplot(data=dt_flow[!ExpID %in% ExpID_to_ignore,], aes(x=factor(X..frequency), y=TXBitrate, fill=factor(X..frequency) ))
g <- g + geom_boxplot()
g <- g + ylab('TXBitrate')
g <- g + xlab('Band ')
fig1  <- ggplotly(g)
fig1
```


```{r flow_channelWidth_TX, echo=FALSE}
dt_flow$tech_channel <- paste0(dt_flow$X..technology, "_", dt_flow$X..channelWidth)
g <- ggplot(data=dt_flow[!ExpID %in% ExpID_to_ignore, ], aes(x=factor(tech_channel), y=TXBitrate, fill=factor(tech_channel) ))
g <- g + geom_boxplot()
g <- g + ylab('TXBitrate')
g <- g + xlab('Per technology (0=ax, 1=n,) and channelBandwidth (ignoring test)')
fig1  <- ggplotly(g)
fig1
```




## RXBitRate


```{r dt_flow_RX, echo=FALSE}

g <- ggplot(data=dt_flow, aes(x=factor(RXBitrate), y=MeanDelay, fill=factor(X..technology), colour=factor(X..numSta) ))
g <- g + geom_boxplot()
g <- g + ylab('RXBitrate per Experiment')
g <- g + xlab('Technology and number of clients within flowID and experiments')
g <- g + facet_wrap(~ExpID, scales="free")
fig2a  <- ggplotly(g)
fig2a
```

**Results:**

* Tech:  
* N. of Clients: 


```{r Flow_RX, echo=FALSE}
g <- ggplot(data=dt_flow[!ExpID %in% ExpID_to_ignore,], aes(x=factor(X..frequency), y=RXBitrate, fill=factor(X..frequency) ))
g <- g + geom_boxplot()
g <- g + ylab('RXBitrate')
g <- g + xlab('Band ')
fig1  <- ggplotly(g)
fig1
```


```{r flow_channelWidth_RX, echo=FALSE}
dt_flow$tech_channel <- paste0(dt_flow$X..technology, "_", dt_flow$X..channelWidth)
g <- ggplot(data=dt_flow[!ExpID %in% ExpID_to_ignore, ], aes(x=factor(tech_channel), y=RXBitrate, fill=factor(tech_channel) ))
g <- g + geom_boxplot()
g <- g + ylab('RXBitrate')
g <- g + xlab('Per technology (0=ax, 1=n,) and channelBandwidth (ignoring test)')
fig1  <- ggplotly(g)
fig1
```



